---
title: "SCD exam"
author: "Manon Grandjean"
date: "4/18/2021"
output: html_document
---

```{r Loading NetLogo plot csv export file - figure 2}
library(tidyverse)
d <- read.csv("NL plots/Happiness plot H1.csv", header = T, skip = 19) # It's supposed to have 16 columns
# nice! I got it to read in all the columns, but only from row 20 in the original csv :) (this is where the data set starts sorta)
# Now I just need to figure out how I 'lengthen' the data frame, because there is kind of 4 parallel sets, one for each breed/color.
```

# Legend: #
Each breed has 4 columns in the data frame associated to it, only the 3 first of which are relevant.
- x, which is the position on the x axis, also the tick count / 'day'. (these are all 0,1,2,3 etc)
- y, which is the value, the population of that given breed and the given tick.
- color, which is the number color assigned in NetLogo. These can be manually changed to chr values red, pink, etc.
- pen.down. just means that the graph is drawing. Not relevant for me. They are all true all of the time.

```{r Split into 4 data frames - figure 2}
dRed <- cbind(d$x, d$y, d$color)
dRed <- as.data.frame(dRed)
colnames(dRed) <- c("x", "y", "color")
dRed$color <- "red"


dPink <- cbind(d$x.1, d$y.1, d$color.1)
dPink <- as.data.frame(dPink)
colnames(dPink) <- c("x", "y", "color")
dPink$color <- "pink"


dDarkblue <- cbind(d$x.2, d$y.2, d$color.2)
dDarkblue <- as.data.frame(dDarkblue)
colnames(dDarkblue) <- c("x", "y", "color")
dDarkblue$color <- "darkblue"


dLightblue <- cbind(d$x.3, d$y.3, d$color.3)
dLightblue <- as.data.frame(dLightblue)
colnames(dLightblue) <- c("x", "y", "color")
dLightblue$color <- "lightblue"

# oh wait, that's not the structure i wanted it in for graphing...
# oh wait maybe i do want the like this and rbind them...
df2 <- rbind(dRed, dPink, dDarkblue, dLightblue)
```


```{r Re-plotting the original NL plot in ggplot - figure 2}
ggplot(df2, aes(x = x, y = y, color = color))+
  geom_line()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

#ggsave("hypothesis1plot.png")
```
What a pretty plot :) Now what can I use it for? What good is it that I now have the ABM data in R?
What things can I do with this data? How can I use it to its fullest potential?

```{r 4.1 figure 3}
#4.1 - figure 3
d <- read.csv("NL plots/meetingandnoreproduction13052021 Happiness plot.csv", header = T, skip = 19)

dRed <- cbind(d$x, d$y, d$color)
dRed <- as.data.frame(dRed)
colnames(dRed) <- c("x", "y", "color")
dRed$color <- "red"


dPink <- cbind(d$x.1, d$y.1, d$color.1)
dPink <- as.data.frame(dPink)
colnames(dPink) <- c("x", "y", "color")
dPink$color <- "pink"


dDarkblue <- cbind(d$x.2, d$y.2, d$color.2)
dDarkblue <- as.data.frame(dDarkblue)
colnames(dDarkblue) <- c("x", "y", "color")
dDarkblue$color <- "darkblue"


dLightblue <- cbind(d$x.3, d$y.3, d$color.3)
dLightblue <- as.data.frame(dLightblue)
colnames(dLightblue) <- c("x", "y", "color")
dLightblue$color <- "lightblue"

df2 <- rbind(dRed, dPink, dDarkblue, dLightblue)

ggplot(df2, aes(x = x, y = y, color = color))+
  geom_line()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()
```


## Aggregating multiple ABM runs

```{r Start by trying to aggrigate the different breeds in the first data set}
# I will start by aggregating these as 1) a proof on concept, 2) to see which data structure is best to do it with.

#trying with df2
#https://www.youtube.com/watch?v=zmiC7X9fUmo
aggregate(y ~ color, df2, FUN = "max")

```
```{r}
((101.5-71.0) / 71.0) * 100
((102.8-77.4) / 77.4) * 100
```



### Processing BehaviorSpace files ###

```{r loading bs data expriment 1}
bs <- read.csv("NL BS outputs/meetingandnoreproduction13052021 hypothesis1-table2.csv", header = T, skip = 6)
```

```{r some descriptive stats and calculations experiment 1}
bs_finaltick <- filter(bs, ticks == 1999)

mean(bs_finaltick$mean..happiness..of.t_reds)
sd(bs_finaltick$mean..happiness..of.t_reds)

mean(bs_finaltick$mean..happiness..of.t_pinks)
sd(bs_finaltick$mean..happiness..of.t_pinks)

mean(bs_finaltick$mean..happiness..of.t_darkblues)
sd(bs_finaltick$mean..happiness..of.t_darkblues)

mean(bs_finaltick$mean..happiness..of.t_lightblues)
sd(bs_finaltick$mean..happiness..of.t_lightblues)

#To calculate the percentage increase:
# First: work out the difference (increase) between the two numbers you are comparing. Then: divide the increase by the 
# original number and multiply the answer by 100. % increase = Increase รท Original Number ร 100.
blue_increase <- mean(bs_finaltick$mean..happiness..of.t_lightblues) - mean(bs_finaltick$mean..happiness..of.t_darkblues)
perc_increace <- (blue_increase / mean(bs_finaltick$mean..happiness..of.t_darkblues)) * 100

red_increase <- mean(bs_finaltick$mean..happiness..of.t_pinks) - mean(bs_finaltick$mean..happiness..of.t_reds)
perc_increace2 <- (red_increase / mean(bs_finaltick$mean..happiness..of.t_reds)) * 100
```


## behavior space experiment 2 ##

```{r load and transform data}
bs2 <- read.csv("NL BS outputs/majorityminority_ex2-table.csv", header = T, skip = 6)

# we want the same number of reds and pinks
rp1 <- filter(bs2, n_red == 1 & n_pink ==1)
rp2 <- filter(bs2, n_red == 2 & n_pink ==2)
rp3 <- filter(bs2, n_red == 3 & n_pink ==3)
rp4 <- filter(bs2, n_red == 4 & n_pink ==4)
rp5 <- filter(bs2, n_red == 5 & n_pink ==5)
rp6 <- filter(bs2, n_red == 6 & n_pink ==6)
rp7 <- filter(bs2, n_red == 7 & n_pink ==7)
rp8 <- filter(bs2, n_red == 8 & n_pink ==8)
rp9 <- filter(bs2, n_red == 9 & n_pink ==9)


rp1$mean..happiness..of.t_reds

# elongate the data
long_bs <- function(d){
  
  dRed <- cbind(d$X.run.number., d$ticks, d$mean..happiness..of.t_reds, d$agreement_reward, d$tolerance_reward, d$intolerance_punishment)
  dRed <- as.data.frame(dRed)
  colnames(dRed) <- c("run.number", "ticks", "mean.happiness", "agreement_reward", "tolerance_reward", "intolerance_punishment")
  dRed$color <- "red"
  
  dPink <- cbind(d$X.run.number., d$ticks, d$mean..happiness..of.t_pinks, d$agreement_reward, d$tolerance_reward, d$intolerance_punishment)
  dPink <- as.data.frame(dPink)
  colnames(dPink) <- c("run.number", "ticks", "mean.happiness", "agreement_reward", "tolerance_reward", "intolerance_punishment")
  dPink$color <- "pink"
  
  dDarkblue <- cbind(d$X.run.number., d$ticks, d$mean..happiness..of.t_darkblues, d$agreement_reward, d$tolerance_reward, d$intolerance_punishment)
  dDarkblue <- as.data.frame(dDarkblue)
  colnames(dDarkblue) <- c("run.number", "ticks", "mean.happiness", "agreement_reward", "tolerance_reward", "intolerance_punishment")
  dDarkblue$color <- "darkblue"
  
  dLightblue <- cbind(d$X.run.number., d$ticks, d$mean..happiness..of.t_lightblues, d$agreement_reward, d$tolerance_reward, d$intolerance_punishment)
  dLightblue <- as.data.frame(dLightblue)
  colnames(dLightblue) <- c("run.number", "ticks", "mean.happiness", "agreement_reward", "tolerance_reward", "intolerance_punishment")
  dLightblue$color <- "lightblue"
  
  dx <- rbind(dRed, dPink, dDarkblue, dLightblue)
  
  return(dx)
}

rp1long <- long_bs(rp1)
rp2long <- long_bs(rp2)
rp3long <- long_bs(rp3)
rp4long <- long_bs(rp4)
rp5long <- long_bs(rp5)
rp6long <- long_bs(rp6)
rp7long <- long_bs(rp7)
rp8long <- long_bs(rp8)
rp9long <- long_bs(rp9)
```

```{r test if it works! yay1!}

ggplot(rp1long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

aggregate(mean.happiness ~ color, rp1long, FUN = "max")
```

```{r big beautiful plot! for experiment 2}
p1 <- ggplot(rp1long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

p2 <- ggplot(rp2long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

p3 <- ggplot(rp3long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

p4 <- ggplot(rp4long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

p5 <- ggplot(rp5long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

p6 <- ggplot(rp6long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

p7 <- ggplot(rp7long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

p8 <- ggplot(rp8long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

p9 <- ggplot(rp9long, aes(x = ticks, y = mean.happiness, color = color))+
  geom_smooth()+
  scale_color_manual(values = c("blue", "deepskyblue", "hotpink", "firebrick3"))+
  #ggtitle("")+
  ylab("Mean happiness")+
  xlab("Time")+
  theme_light()

ggpubr::ggarrange(p1, p2, p3, p4, p5, p6, p7, p8, p9, 
                  labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9"),
                  ncol = 3, nrow = 3,
                  common.legend = TRUE)

ggsave("minorityplot9.png", width = 10)
```


